A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-08-21 tex/paper.tex
plotting available

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
the advantage of being trivially fast, but each index relationship is only informative over a 
limited region of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
used (particularly in the spectral dimension), thereby achieving higher precision even for data 
with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
a spectral synthesis front-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
practice, this often requires an effort focused on a pre-selected subsample of the data.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``front-end" that generates a model spectrum.  In our adopted approach, the model 
spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
\begin{equation} \label{eqn:interp}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesirable level of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
$\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the interpolated spectrum with the 
corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
%% \todo{tests regarding interpolation accuracy}.

@@

The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
straightforward operations demanded by practical astronomical and computing issues.  If the data 
were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
the physical parameters.  This concern is treated externally to any modeling procedure, most often 
by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
``normalization" implicitly assumes that there is no relevant information content in the spectral 
shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
molecular bands for cool stars).

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
completely independent samples of the true spectrum.  In that case, a difference between an 
observed and modeled spectral feature will create a correlated residual that spans multiple 
pixels.  This can be demonstrated practically through the autocorrelation of $\vR$: a slight model 
mismatch will produce correlated residuals over a characteristic scale similar to the instrumental
broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$4 
pixel scale, corresponding to the 3\,km s$^{-1}$ width of $\mathcal{F}_v^{\rm inst}$.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{correlation_panel.pdf}
  \caption{\textbf{Top}: the big spectrum. \textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.   
  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariant structure should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ include an amplitude ($a_g$) and 
scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w_{ij}^{\rm g} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matern} demonstrates that random draws from such a 
kernel readily produce correlated structure similar to those seen in a typical residual spectrum.

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters. There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$). This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix. The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals. The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
\label{fig:matern}
\end{center}
\end{figure*}

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual could be described as
\begin{equation}
\vR = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
residual regions can be written
\begin{equation} \label{eqn:kregion}
  \mathcal{K}^{l,k}_{ij} = \frac{a_k^2}{2\pi\sigma_k^2} \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
\end{equation}
such that the full local covariance kernel composed of a sequence of residual regions is the linear 
combination,
\begin{equation} \label{eqn:klocal}
\Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{l,k}_{ij},
\end{equation}
with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure that a sparse covariance 
matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:region} shows how this non-stationary Gaussian 
kernel generates a localized region of enhanced variance that successfully mimics the kind of 
residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
they might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
the probabilistic framework (as opposed to manual or threshold-based clipping or masking).

\begin{figure*}[!t]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}). The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals. These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix. The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix. The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
\label{fig:region}
\end{center}
\end{figure*}

@@

%\begin{figure*}[!t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}). The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals. These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix. The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix. The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
%\label{fig:region}
%\end{center}
%\end{figure*}

@@

We assume uniform priors on the global covariance parameters.  For the local covariance kernels, we 
adopt flat priors for the amplitudes and means, \{$a_k$, $\mu_k$\}, and construct a specialized 
prior for the widths, \{$\sigma_k$\}, that is flat when $\sigma_k < \sigma_v$ and has a rapid 
(Gaussian) taper to zero for larger widths.  This form is designed to prevent the local 
kernels from diffusing to large $\sigma_k$ and low $a_k$, since such behavior is better treated by 
the global kernel.

@@

\item Decide whether to instantiate any local (non-stationary) covariance kernels.  We adopt an 
iterative thresholding criterion: if a local patch of the residual signal is at least $3\times$ the 
amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this triggers the 
addition of a local kernel.\footnote{There is no practical reason to delete local kernels.  If the 
parameters have changed such that a given local kernel is no longer required, that kernel amplitude 
will be driven towards zero and represent a negligible contribution to $\vC$; in effect, the model 
will act as if the kernel were deleted automatically.}

\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
$\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
$\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
\end{enumerate}

@@

After burn-in, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
inter-chain variances to assess how the inference of the posterior would be improved if the chains 
were to be run for an infinite number of iterations.  The Gibbs sampler is eventually stopped if 
$\hat{R} \, < \, 1.1$, and the chains appear converged in a visual inspection.

@@

@@

>>>>

%Chebyshev coefficients RULE!
\newcommand{\cc}[2]{c_{#2}^{(#1)}} %echelle order first, then degree coefficient

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
the advantage of being trivially fast, but each index relationship is only informative over a 
limited region of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
used (particularly in the spectral dimension), thereby achieving higher precision even for data 
with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
a spectral synthesis back-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
practice, this often requires an effort focused on a pre-selected subsample of the data.

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model 
spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
\begin{equation} \label{eqn:interp}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesirable level of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
$\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the interpolated spectrum with the 
corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation
technique is found to be accurate within a few percent per full-resolution
model pixel and $\lesssim 1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
this pre-interpolation could be avoided if the spectral library was computed
over a refined grid (with a substantial up-front computational investment); but
for the time being, we can empirically propagate these interpolation
uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
straightforward operations demanded by practical astronomical and computing issues.  If the data 
were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
the physical parameters.  This concern is usually treated externally to any modeling procedure, most often 
by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
``normalization" implicitly assumes that there is no relevant information content in the spectral 
shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
molecular bands for cool stars).

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
completely independent samples of the true spectrum.  In that case, a difference between an 
observed and modeled spectral feature will create a correlated residual that spans multiple 
pixels.  This can be demonstrated practically through the autocorrelation of $\vR$: a slight model 
mismatch will produce correlated residuals over a characteristic scale similar to the instrumental
broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{correlation_panel.pdf}
  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
  parameters drawn from the posterior distribution, and the residuals that
  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
  to show the mildly covariant structure produced by slight mismatch between
  the data and model spectra.   \textbf{Right} The autocorrelation of the
residual sequence shown at left. Notice that there is significant correlation
for offsets of $\lesssim 5$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariant structure should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
residuals, 
\begin{equation}
  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
kernel with $\nu = 3/2$,
\begin{equation}
  \Kglobal_{ij}(\vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
where the (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ include an amplitude ($a_g$) and 
scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w_{ij}^{\rm g} \,(r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matrix} demonstrates that random draws from such a 
kernel readily produce correlated structure similar to those seen in a typical residual spectrum.

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/matrix_compilation.pdf}
\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals. \textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region. Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix. \textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column. At the bottom of each panel is the envelope formed by 50 random draws (blue) overlaid with the actual residuals from the gray region. \textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum. \textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA. \textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals. Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines. The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit. Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not. All parameters of the covariance kernels are determined self-consistently along with the stellar parameters. }
\label{fig:matrix}
\end{center}
\end{figure*}

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual could be described as
\begin{equation}
\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
residual regions can be written
\begin{equation} \label{eqn:kregion}
  \mathcal{K}^{l,k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
\end{equation}
such that the full local covariance kernel composed of a sequence of residual regions is the linear 
combination,
\begin{equation} \label{eqn:klocal}
\Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{l,k}_{ij},
\end{equation}
with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:matrix} shows how this non-stationary Gaussian 
kernel generates a localized region of enhanced variance that successfully mimics the kind of 
residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
they might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
the probabilistic framework (as opposed to manual or threshold-based clipping or masking).

@@

@@

We assume uniform priors on the global covariance parameters.  For the local
covariance kernels, we adopt flat priors for the amplitudes and means, \{$a_k$,
$\mu_k$\}, and construct a specialized prior for the widths, \{$\sigma_k$\},
that is flat below the instrumental width (when $\sigma_k < \sigma_v$) and has
a rapid (logistic) taper to zero for larger widths.  This form is designed to
prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
since such behavior is better treated by the global kernel.

@@

\item If any local (non-stationary) covariance kernels have been instantiated (see below), sample in their hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
$\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
$\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
\end{enumerate}

@@

Local kernels are instantiated by the following procedure:  An ``average''
residual spectrum is created by combining ~500 spectra residuals stored from a
burn-in period using only global covariance kernels (thinned to account for
autocorrelation of the Markov chain). Then, using an iterative thresholding
criterion similar in procedure to
``sigma clipping,''\footnote{Although the procedure of identifying outliers is
  similar to sigma clipping, the inference on the stellar parameters is
  emphatically \emph{not} the same as sigma clipping, since we now determine
the weights of these spectral outliers inside of a probabilistic framework, rather than
directly setting their weights to zero as in sigma clipping.} we identify the largest outlier
pixel not already covered by a local kernel, instantiate a local kernel at its location, and repeat until there are no pixels above a certain
threshold left.\footnote{There is no practical
  reason to delete local kernels once instantiated.  If the parameters have
  changed such that a given local kernel is no longer required, that kernel
  amplitude will be driven towards zero and represent a negligible contribution
  to $\vC$; in effect, the model will act as if the kernel were deleted
automatically.} Alternative local kernel instantiation schemes,
such as re-evaluating the location of kernels with each Gibbs sampler
iteration, have yielded similar results to the thresholding approach. However, this approach most consistently converges and only has a modest computational overhead. We choose four times the standard deviation of the average residual spectrum as our threshold. Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which in turn slightly reduces the amplitude of the global kernel. Taken to the logical conclusion, if the threshold was set low enough then there would be a local kernel for each spectral line and no global kernel would be required. We find that the final stellar posterior is relatively insensitive to the choice of threshold level as long as it is set low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$). Once all of the local kernels have been instantiated, the Gibbs sampler is run for another period of burn-in.

After burn-in is complete, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
inter-chain variances to assess how the inference of the posterior would be improved if the chains 
were to be run for an infinite number of iterations.  The Gibbs sampler is eventually stopped if 
$\hat{R} \, < \, 1.1$, and the chains appear converged in a visual inspection.

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_Kurucz_logg.pdf}
\caption{Kurucz WASP-14.}
\label{fig:residuals_Kurucz}
\end{center}
\end{figure*}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6516.8 & 12.6 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.62 & 0.02 $\kms$ \\
$\log_{10} \Omega$ & -12.72 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\cc{22}{0}$ & & \\
$\cc{22}{1}$ & & \\
$\cc{22}{2}$ & & \\
$\cc{22}{3}$ & & \\
$b_{22}$ & & \\
$a_{{\rm g}, 22}$ & & \\
$l_{22}$ & & \\
\cutinhead{Order 23 nuisance parameters}
$\cc{23}{0}$ & & \\
$\cc{23}{1}$ & & \\
$\cc{23}{2}$ & & \\
$\cc{23}{3}$ & & \\
$b_{23}$ & & \\
$a_{{\rm g}, 23}$ & & \\
$l_{23}$ & & \\
\cutinhead{Order 24 nuisance parameters}
$\cc{24}{0}$ & & \\
$\cc{24}{1}$ & & \\
$\cc{24}{2}$ & & \\
$\cc{24}{3}$ & & \\
$b_{24}$ & & \\
$a_{{\rm g}, 24}$ & & \\
$l_{24}$ & & \\
\enddata
\end{deluxetable}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics{figs/residuals_PHOENIX_logg.pdf}
\caption{Kurucz WASP-14.}
\label{fig:residuals_PHOENIX}
\end{center}
\end{figure*}

\begin{deluxetable}{lr@{$\pm$}l} 
\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
\startdata
\cutinhead{Stellar parameters}
$T_{\rm eff}$ & 6516.8 & 12.6 K \\
$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
$\Z$ & -0.27 & 0.01 dex \\
$v \sin i$ & 4.29 & 0.05 $\kms$\\
$v_z$ & -4.62 & 0.02 $\kms$ \\
$\log_{10} \Omega$ & -12.72 & 0.0004 \\
\cutinhead{Order 22 nuisance parameters}
$\cc{22}{0}$ & & \\
$\cc{22}{1}$ & & \\
$\cc{22}{2}$ & & \\
$\cc{22}{3}$ & & \\
$b_{22}$ & & \\
$a_{{\rm g}, 22}$ & & \\
$l_{22}$ & & \\
\cutinhead{Order 23 nuisance parameters}
$\cc{23}{0}$ & & \\
$\cc{23}{1}$ & & \\
$\cc{23}{2}$ & & \\
$\cc{23}{3}$ & & \\
$b_{23}$ & & \\
$a_{{\rm g}, 23}$ & & \\
$l_{23}$ & & \\
\cutinhead{Order 24 nuisance parameters}
$\cc{24}{0}$ & & \\
$\cc{24}{1}$ & & \\
$\cc{24}{2}$ & & \\
$\cc{24}{3}$ & & \\
$b_{24}$ & & \\
$a_{{\rm g}, 24}$ & & \\
$l_{24}$ & & \\
\enddata
\end{deluxetable}
